Supplementary Material for: A Rough Guide to Pre-processing High-Frequency Animal Tracking Data

### 2021-02-03 06_bat_data.Rmd
change title
style file

# Processing Egyptian fruit bat tracks

@@

```{r read_bat_data}
# prepare the connection
con <- dbConnect(drv = SQLite(), 
                 dbname = "data/Three_example_bats.sql")

@@

```{r echo=FALSE}
# plot bat data to examine for errors
fig_bat_raw <-
  ggplot(data)+
  geom_path(aes(X, Y,
                group = TAG,
                colour = TAG),
            show.legend = FALSE)+
  scale_colour_brewer(palette = "Dark2")+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  coord_sf(crs = 2039)
```

```{r echo=FALSE}
# save to figures
ggsave(fig_bat_raw,
       filename = "figures/fig_bat_raw.png",
       dpi = 300,
       width = 6, height = 4)
```

@@

```{r}
# filter for SD <= 20
# here, reassignment is necessary as rows are being removed
# the atl_filter_covariates function could have been used here
data_split <- lapply(data_split, function(dt) {

  dt <- atl_filter_covariates(
    data = dt,
    filters = c("SD <= 20",
                "NBS > 3")

  )
})
```

@@

```{r echo=FALSE}
# now plot over raw data
fig_bat_filter_cov <- 
  ggplot()+
  geom_point(data = data[!data_split, 
                         on = c("TAG", "X", "Y", "TIME")],
             aes(X, Y),
             col = "grey",
             shape = 4)+
  geom_path(data = data_split,
             aes(X, Y,
                 col = TAG),
             size = 0.3,
             show.legend = FALSE)+
  scale_colour_brewer(palette = "Dark2")+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  coord_sf(crs = 2039,
           xlim = range(data_split$X),
           ylim = range(data_split$Y))
```

```{r echo=FALSE}
# save to figures
ggsave(fig_bat_filter_cov,
       filename = "figures/fig_bat_filter_cov.png",
       dpi = 300,
       width = 4, height = 6)
```

@@

    dt[, `:=`(speed_in = atl_get_speed(dt, 
                                       x = "X", y = "Y", 
                                       time = "time",
                                       type = "in"),
              speed_out = atl_get_speed(dt, 
                                       x = "X", y = "Y", 
                                       time = "time",
                                       type = "out"))]
  })
)
```

@@

  dt <- atl_filter_covariates(data = dt,
                              filters = c("speed_in <= 20",
                                          "speed_out <= 20"))
})
```

@@

```{r echo=FALSE}
# now plot over raw data
fig_bat_filter_speed <-
  ggplot()+
  geom_point(data = data[!data_split, 
                         on = c("TAG", "X", "Y", "TIME")],
             aes(X, Y),
             col = "grey",
             shape = 4)+
  geom_path(data = data_split,
             aes(X, Y,
                 col = TAG),
             size = 0.3,
             show.legend = FALSE)+
  scale_colour_brewer(palette = "Dark2")+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  coord_sf(crs = 2039,
           xlim = range(data_split$X),
           ylim = range(data_split$Y))
```

```{r echo=FALSE}
# save to figures
ggsave(fig_bat_filter_speed,
       filename = "figures/fig_bat_filter_speed.png",
       dpi = 300,
       width = 4, height = 6)
```

@@

  lapply(X = data_smooth, 
         FUN = atl_median_smooth,
         time = "time", moving_window = 5)
)
```

@@

```{r echo=FALSE}
# plot the raw, filtered, and smoothed data
fig_bat_smooth <- 
  ggplot()+
  geom_point(data = data_split,
             aes(X, Y),
             col = "grey",
             shape = 4,
             show.legend = FALSE)+
  geom_path(data = data_smooth,
            aes(X, Y,
                col = TAG),
            lwd = 0.3,
            show.legend = FALSE)+
  scale_colour_brewer(palette = "Dark2")+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  coord_sf(crs = 2039,
           xlim = range(data_smooth$X),
           ylim = range(data_smooth$Y))
```

```{r echo=FALSE}
# save to figures
ggsave(fig_bat_smooth,
       filename = "figures/fig_bat_smooth.png",
       dpi = 300,
       width = 4, height = 6)
```

@@

  dt <- merge(dt, 
              dt_recurse[, c("coordIdx", "resTime")], 
              by = c("coordIdx"))

@@

```{r echo=FALSE, eval=FALSE}
# not shown or mentioned
fig_bat_res_timeseries <-
  ggplot()+
  geom_path(data = data_residence,
            aes(ts, resTime,
                colour = TAG),
            show.legend = F)+
  scale_color_brewer(palette = "Dark2")+
  scale_x_datetime()+
  ggthemes::theme_few()+
  labs(x = NULL,
       y = "residence time (mins)")
```

```{r echo=FALSE}
# not shown or mentioned
# save to figures
ggsave(fig_bat_res_timeseries,
       filename = "figures/fig_bat_res_timeseries.png",
       dpi = 300,
       width = 6, height = 3)
```

@@

```{r}
# add an id column
data_residence[, `:=`(id = TAG,
                      x = X, y = Y)]

@@

```{r}
# segment into residence patches
data_patches <- lapply(data_residence, atl_res_patch,
                       buffer_radius = 25)
```

@@

```{r}
# get data spatials
data_spatials <- lapply(data_patches, atl_patch_summary,
                        which_data = "spatial",
                        buffer_radius = 25)

@@

```{r}
st_write(data_spatials,
         dsn = "data/data_bat_residence_patches.gpkg")
```

@@

```{r}
data_clean <- fwrite(rbindlist(data_smooth),
                     file = "data/data_bat_smooth.csv")
```

@@

# write
fwrite(patch_summary,
       "data/data_bat_patch_summary.csv")
```

>>>>

# Processing Egyptian Fruit Bat Tracks

@@

```{r read_bat_data}
# prepare the connection
con <- dbConnect(
  drv = SQLite(),
  dbname = "data/Three_example_bats.sql"
)

@@

```{r echo=FALSE}
# plot bat data to examine for errors
fig_bat_raw <-
  ggplot(data) +
  geom_path(aes(X, Y,
    group = TAG,
    colour = TAG
  ),
  show.legend = FALSE
  ) +
  scale_colour_brewer(palette = "Dark2") +
  ggspatial::annotation_scale(location = "br") +
  ggthemes::theme_few() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  coord_sf(crs = 2039)
```

```{r echo=FALSE}
# save to figures
ggsave(fig_bat_raw,
  filename = "figures/fig_bat_raw.png",
  dpi = 300,
  width = 6, height = 4
)
```

@@

```{r}
# filter for SD <= 20
# here, reassignment is necessary as rows are being removed
# the atl_filter_covariates function could have been used here
data_split <- lapply(data_split, function(dt) {
  dt <- atl_filter_covariates(
    data = dt,
    filters = c(
      "SD <= 20",
      "NBS > 3"
    )
  )
})
```

@@

```{r echo=FALSE}
# now plot over raw data
fig_bat_filter_cov <-
  ggplot() +
  geom_point(
    data = data[!data_split,
      on = c("TAG", "X", "Y", "TIME")
    ],
    aes(X, Y),
    col = "grey",
    shape = 4
  ) +
  geom_path(
    data = data_split,
    aes(X, Y,
      col = TAG
    ),
    size = 0.3,
    show.legend = FALSE
  ) +
  scale_colour_brewer(palette = "Dark2") +
  ggspatial::annotation_scale(location = "br") +
  ggthemes::theme_few() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  coord_sf(
    crs = 2039,
    xlim = range(data_split$X),
    ylim = range(data_split$Y)
  )
```

```{r echo=FALSE}
# save to figures
ggsave(fig_bat_filter_cov,
  filename = "figures/fig_bat_filter_cov.png",
  dpi = 300,
  width = 4, height = 6
)
```

@@

    dt[, `:=`(
      speed_in = atl_get_speed(dt,
        x = "X", y = "Y",
        time = "time",
        type = "in"
      ),
      speed_out = atl_get_speed(dt,
        x = "X", y = "Y",
        time = "time",
        type = "out"
      )
    )]
  })
)
```

@@

  dt <- atl_filter_covariates(
    data = dt,
    filters = c(
      "speed_in <= 20",
      "speed_out <= 20"
    )
  )
})
```

@@

```{r echo=FALSE}
# now plot over raw data
fig_bat_filter_speed <-
  ggplot() +
  geom_point(
    data = data[!data_split,
      on = c("TAG", "X", "Y", "TIME")
    ],
    aes(X, Y),
    col = "grey",
    shape = 4
  ) +
  geom_path(
    data = data_split,
    aes(X, Y,
      col = TAG
    ),
    size = 0.3,
    show.legend = FALSE
  ) +
  scale_colour_brewer(palette = "Dark2") +
  ggspatial::annotation_scale(location = "br") +
  ggthemes::theme_few() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  coord_sf(
    crs = 2039,
    xlim = range(data_split$X),
    ylim = range(data_split$Y)
  )
```

```{r echo=FALSE}
# save to figures
ggsave(fig_bat_filter_speed,
  filename = "figures/fig_bat_filter_speed.png",
  dpi = 300,
  width = 4, height = 6
)
```

@@

  lapply(
    X = data_smooth,
    FUN = atl_median_smooth,
    time = "time", moving_window = 5
  )
)
```

@@

```{r echo=FALSE}
# plot the raw, filtered, and smoothed data
fig_bat_smooth <-
  ggplot() +
  geom_point(
    data = data_split,
    aes(X, Y),
    col = "grey",
    shape = 4,
    show.legend = FALSE
  ) +
  geom_path(
    data = data_smooth,
    aes(X, Y,
      col = TAG
    ),
    lwd = 0.3,
    show.legend = FALSE
  ) +
  scale_colour_brewer(palette = "Dark2") +
  ggspatial::annotation_scale(location = "br") +
  ggthemes::theme_few() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  coord_sf(
    crs = 2039,
    xlim = range(data_smooth$X),
    ylim = range(data_smooth$Y)
  )
```

```{r echo=FALSE}
# save to figures
ggsave(fig_bat_smooth,
  filename = "figures/fig_bat_smooth.png",
  dpi = 300,
  width = 4, height = 6
)
```

@@

  dt <- merge(dt,
    dt_recurse[, c("coordIdx", "resTime")],
    by = c("coordIdx")
  )

@@

```{r echo=FALSE, eval=FALSE}
# not shown or mentioned
fig_bat_res_timeseries <-
  ggplot() +
  geom_path(
    data = data_residence,
    aes(ts, resTime,
      colour = TAG
    ),
    show.legend = F
  ) +
  scale_color_brewer(palette = "Dark2") +
  scale_x_datetime() +
  ggthemes::theme_few() +
  labs(
    x = NULL,
    y = "residence time (mins)"
  )
```

```{r echo=FALSE}
# not shown or mentioned
# save to figures
ggsave(fig_bat_res_timeseries,
  filename = "figures/fig_bat_res_timeseries.png",
  dpi = 300,
  width = 6, height = 3
)
```

@@

```{r}
# add an id column
data_residence[, `:=`(
  id = TAG,
  x = X, y = Y
)]

@@

```{r}
# segment into residence patches
data_patches <- lapply(data_residence, atl_res_patch,
  buffer_radius = 25
)
```

@@

```{r}
# get data spatials
data_spatials <- lapply(data_patches, atl_patch_summary,
  which_data = "spatial",
  buffer_radius = 25
)

@@

```{r}
st_write(data_spatials,
  dsn = "data/data_bat_residence_patches.gpkg"
)
```

@@

```{r}
data_clean <- fwrite(rbindlist(data_smooth),
  file = "data/data_bat_smooth.csv"
)
```

@@

# write
fwrite(
  patch_summary,
  "data/data_bat_patch_summary.csv"
)
```
